---
title: "p8105_hw2_tw2838"
author: "Tianyou Wang"
date: "2022-10-05"
output: html_document
---

Set up and load the libraries.
```{r warning = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


# **Problem #1 NYC Transit data**

Import and clean the NYC Transit csv file.

```{r}
nyc_df = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, station_longitude,
         route1:route11, entry, vending, entrance_type, ada) %>%
  mutate(entry = ifelse(entry == "YES", 1, 0))
```
After we cleaned the data, this dataset contains `r nrow(nyc_df)` observations, and it is charactrized by `r ncol(nyc_df)` variables.

# **Problem #2 Mr.Trash Wheel**

Import and clean the Mr. Trash Wheel excel sheet.

```{r warning = FALSE}
tw_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = 1, skip = 1) %>%
  janitor::clean_names() %>% 
  select(-x15, -x16) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>% 
  mutate(type = "Trash Wheel") %>%
  drop_na(dumpster) %>%
  view()
```

Import and clean the Professor Trash Wheel sheet.

```{r warning = FALSE}
ptw_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = 2, skip = 1) %>%
  janitor::clean_names() %>%
  mutate(type = "Professor Trash Wheel") %>%
  mutate(sports_balls = "NA") %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(year = as.character(year)) %>%
  relocate(sports_balls, .after = chip_bags) %>%
  drop_na(dumpster) %>%
  view()
```

Combining the Mr. Trash Wheel and Professor Trash Wheel into a single tidy dataset.

```{r warning = FALSE}
com_df = bind_rows(tw_df, ptw_df) %>%
  relocate(type, everything()) %>%
view()
```


In Mr.Trash Wheel, we have `r nrow(tw_df)` observations, and it is charactrized by `r ncol(tw_df)` variables. We can see the example of the first 5 observations by using `print(tw_df, n=5)`. 

```{r warning = FALSE}
print(tw_df, n=5)
```


Similiarly, we have `r nrow(ptw_df)` observations in the Professional Trash Wheel datasete which is also charactrized by `r ncol(ptw_df)` variables. We can check the first 5 observations by using `print(tw_df, n=5)`.
 
```{r warning = FALSE}
print(ptw_df, n=5)
```


The total weight of trash collected by Professor Trash Wheel is `r sum(ptw_df$weight_tons)` tons.

The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(tw_df %>% filter(year == 2020) %>% pull(sports_balls))`.



# **Problem #3 FiveThirtyEight **

Import and clean the dataset: pols-month.csv

```{r warning=FALSE}
pm_df = read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), sep = '-') %>%
  mutate(month = month.name[as.integer(month)]) %>%
  mutate(president = ifelse(prez_gop == 0, "dem", "gop")) %>%
  select(-prez_gop, -prez_dem, -day)
```

Import and clean the dataset: snp.csv

```{r warning=FALSE}
snp_df = read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("month", "day", "year"), sep = '/') %>%
  mutate(month = month.name[as.integer(month)]) %>%
  relocate(year, month, everything()) %>%
  select(-day) %>%
  mutate(year = as.integer(year)) %>%
  mutate(year = ifelse(year > 15 , year + 1900, year + 2000)) %>%
  mutate(year = as.character(year)) %>%
  arrange(year, month)
```

Import and clean the dataset: unemployment.csv

```{r warning=FALSE}
une_df = read_csv("./data/fivethirtyeight_datasets/unemployment.csv") %>% 
  janitor::clean_names() %>% 
  mutate(year = as.character(year)) %>%
  pivot_longer(jan:dec, 
    names_to = "month",
    values_to = "unemployment"
  )
```

Combine snp and pols
```{r}
comb1_df = left_join(pm_df, snp_df, by = c("year", "month")) 
```

Combine comb1 and unemployment
```{r}
comb2_df = left_join(comb1_df,une_df, by = c("year", "month")) 
```

In the final combined dataset, we have data collected for all months from `min(comb2_df$year)` to `max(comb2_df$year)`. 

It is charactrized by `r ncol(comb2_df)` variables. We can see the example of the first 5 observations by using `print(tw_df, n=5)`. 

```{r}
print(tw_df, n=5)
```
