---
title: "p8105_hw2_tw2838"
author: "Tianyou Wang"
date: "2022-10-05"
output: html_document
---

Set up and load the libraries.
```{r warning = FALSE}
library(tidyverse)
library(readxl)
library(haven)
```


# **Problem #1**



# **Problem #2 Mr.Trash Wheel**

Import and clean the Mr. Trash Wheel excel sheet.

```{r warning = FALSE}
tw_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = 1, skip = 1) %>%
  janitor::clean_names() %>% 
  select(-x15, -x16) %>%
  mutate(sports_balls = as.integer(sports_balls)) %>% 
  mutate(type = "Trash Wheel") %>%
  drop_na(dumpster) %>%
  view()
```

Import and clean the Professor Trash Wheel sheet.

```{r warning = FALSE}
ptw_df = read_excel("./data/Trash_Wheel_Collection_Data.xlsx", sheet = 2, skip = 1) %>%
  janitor::clean_names() %>%
  mutate(type = "Professor Trash Wheel") %>%
  mutate(sports_balls = "NA") %>%
  mutate(sports_balls = as.integer(sports_balls)) %>%
  mutate(year = as.character(year)) %>%
  relocate(sports_balls, .after = chip_bags) %>%
  drop_na(dumpster) %>%
  view()
```

Combining the Mr. Trash Wheel and Professor Trash Wheel into a single tidy dataset.

```{r warning = FALSE}
com_df = bind_rows(tw_df, ptw_df) %>%
  relocate(type, everything()) %>%
view()
```


In Mr.Trash Wheel, we have `r nrow(tw_df)` observations, and it is charactrized by `r ncol(tw_df)` variables. We can see the example of the first 5 observations by using `print(tw_df, n=5)`. 

```{r warning = FALSE}
print(tw_df, n=5)
```


Similiarly, we have `r nrow(ptw_df)` observations in the Professional Trash Wheel datasete which is also charactrized by `r ncol(ptw_df)` variables. We can check the first 5 observations by using `print(tw_df, n=5)`.
 
```{r warning = FALSE}
print(ptw_df, n=5)
```


The total weight of trash collected by Professor Trash Wheel is `r sum(ptw_df$weight_tons)` tons.

The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r sum(tw_df %>% filter(year == 2020) %>% pull(sports_balls))`.



# **Problem #3**